A Multi-Fidelity Framework for Wildland Fire Behavior Simulations over Complex Terrain

### 2021-01-24 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

% Contact information of the corresponding author
\corres{Correspondence: marcos.vanella@nist.gov}

@@

\subsubsection{Unresolved Boundary Fluxes}
\label{sec:boundflx}
In this section, the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries are discussed.  Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main point of this paper).

@@

There are various ways of simulating fire in vegetation which can depend on the fire behavior of interest. For example, for a fire spreading through surface vegetation (e.g., grass) the fire behavior differs depending on the orientation of the local wind relative to the direction of fire spread. For head fires, the wind flow and fire spread are in the same direction. Back fires spread into the wind. Flank fires spread approximately orthogonal to wind. For a surface fire spreading from a point in a uniform vegetation, the head fire portion of the resulting fireline has the largest flame base (easiest to resolve) and is characterized by fastest spread and flames with larger lengths that tilt toward the unburned vegetation. Backfires have the slowest spread and smallest flame base (require finer grid cells than the head fire) and have flames with smaller lengths that tilt away from the unburned vegetation. The roles of radiant and convective heat transfer differ along the fireline. Heat gain b
\begin{enumerate}
\item {\bf Particle Model:} The vegetation (surface and raised) is represented by a collection of Lagrangian particles that are heated via convection and radiation. This model, with sufficient grid resolution, is appropriate for head, back, and flank surface fires, as well as fire through raised vegetation (e.g., trees). Heat transfer in the volume containing the vegetation is modeled in all three directions. This model is appropriate for grid resolutions of the order of 1~m or less, depending on the size of the flame base and properties of the vegetation (e.g.,~\cite{Perez-Ramirez:FT2017}).
\item {\bf Boundary Fuel Model:} Surface vegetation has its own grid and is modeled like a porous solid with a thickness equal to the height of the vegetation. This model was designed for head fire spread in surface vegetation based on the assumption that heat transfer in the fuel bed is dominated by radiation from the overhead flame and therefore in the vertical direction.  In the implementation here, the height of the surface vegetation is assumed to unresolved on the grid. The appropriate gas-phase grid resolution of the order of 1~m to 10~m.
\item {\bf Level Set Method:} The fire front of a surface fire propagates using purely empirical rules in a level set method. Thermal degradation of the surface vegetation is not modeled. More than one implementation of this method is possible, largely differentiated by how the wind and fire-atmosphere interaction is modeled. The simpler implementations of this model can use grid resolutions coarser, 10 m and greater, than the more physics-based particle and boundary fuel models. The level set model can be used for firespread in the surface vegetation along with the particle model for fire behavior in raised vegetation. 
\end{enumerate}
The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels and the fire spread rate is {\em predicted} by the model. The Level Set Method relies on a set of predetermined spread rates for different types of vegetation and wind speeds. The pyrolysis model, developed specifically for vegetation~\cite{Porterie:2006,Morvan:CF2004,Houssami:2016}, consists of three reactions:

@@

In FDS, the level set spread model can be implemented in the following ways:
\begin{enumerate}
\item Only the level set simulation is performed, with a constant and uniform specified wind and slope. The wind is not affected by the terrain, and there is no fire.
\item The wind field is established over the terrain, but it is ``frozen'' when the fire ignites.
\item The wind field follows the terrain, but there is no actual fire in the simulation, just front-tracking.  The level set evolves continuously in time with the flow field.
\item The wind and fire are fully-coupled and the resulting wind values are used in the head fire spread rate formula. When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
\item The wind and fire are fully-coupled in the gas-phased, but the head fire spread rate is not influenced by the wind speed. 
\end{enumerate}
When fully coupled to the CFD model, the level set function acts as an igniter as it reaches a new surface cell. The cell burns for a pre-determined amount of time. For cases where the depth of the fireline is less than the size of a grid cell, the burning time is extended while the burning rate is decreased so as to maintain conservation of fuel mass.

@@

\begin{table}[ht]
\begin{center}
\caption[Measured properties for the CSIRO Grassland Fire cases]{Measured properties for the CSIRO Grassland Fire cases~\cite{Cheney:IJWF1993}.}
\label{Properties_Grasses}
\begin{tabular}{|l|c|c|c|}
\hline
Property                        & Units        & Case C064 & Case F19    \\ \hline \hline
Wind Speed                      & m/s          & 4.6       & 4.8         \\ \hline
Ambient Temperature             & $^\circ$C    & 32.       & 34          \\ \hline
Surface Area to Volume Ratio    & m$^{-1}$     & 9770      & 12240       \\ \hline
Grass Height                    & m            & 0.21      & 0.51        \\ \hline
Bulk Mass per Unit Area         & kg/m$^2$     & 0.283     & 0.313       \\ \hline
Moisture Fraction               & \%           & 6.3       & 5.8         \\ \hline
Measured RoS                    & m/s          & 1.0       & 1.5         \\ \hline
Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s           & 1.4, 1.3, 1.4           & 1.1, 1.2, 1.2          \\ \hline
Calc'd RoS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s      & 1.5, 1.4, 1.7           & 1.3, 1.3, 1.3           \\ \hline
Calc'd RoS, Level Set Method (5 m, 10 m, 20 m resolution) & m/s          & 1.1, 1.1, 1.2          & 0.4, 0.5, 0.8          \\ \hline
\end{tabular}
\end{center}
\end{table}

>>>>

% Contact information of the corresponding author
\corres{Correspondence: marcos.vanella@nist.gov, kevin.mcgrattan@nist.gov}

@@

\subsubsection{Unresolved Boundary Fluxes}
\label{sec:boundflx}
In this section, the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries are discussed.  Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main focus of this paper).

@@

There are various types of wildland fire spread models to choose from, and the choice depends on the desired level of physical fidelity and the available computing resources. The choice also depends on the nature of the fire. For example, for surface vegetation like dry grass, the fire behavior depends on the orientation of the local wind relative to the direction of fire spread. For head fires, the wind flow and fire spread are in the same direction. Back fires spread against the wind; flank fires spread laterally. Typically, the head fire exhibits the widest front width, most rapid spread, and longest flame lengths that tilt toward the virgin vegetation. Back fires exhibit the least rapid spread, most narrow front width, and shortest flame lengths that tilt away from the virgin vegetation. Of these three types of fire, the head fire is least demanding in terms of grid resolution; the back fire the most.
\begin{enumerate}
\item {\bf Particle Model:} The vegetation (surface and raised) is represented by a collection of Lagrangian particles that are heated via convection and radiation. This model, with sufficient grid resolution, is appropriate for head, back, and flank surface fires, as well as fire through raised vegetation (e.g., trees). Heat transfer in the volume containing the vegetation is modeled in all three directions. This model is appropriate for grid resolutions of the order of 1~m or less, depending on the size of the flame base and properties of the vegetation (e.g.,~\cite{Perez-Ramirez:FT2017}).
\item {\bf Boundary Fuel Model:} Surface vegetation has its own grid and is modeled like a porous solid with a thickness equal to the height of the vegetation. This model was designed for head fire spread in surface vegetation based on the assumption that heat transfer in the fuel bed is dominated by radiation from the overhead flame and therefore in the vertical direction.  In the implementation here, the height of the surface vegetation is assumed to be unresolved on the grid. The appropriate gas-phase grid resolution of the order of 1~m to 10~m.
\item {\bf Level Set Method:} The fire front of a surface fire propagates using purely empirical rules in a level set method. Thermal degradation of the surface vegetation is not modeled. More than one implementation of this method is possible, largely differentiated by how the wind and fire-atmosphere interaction is modeled. The simpler implementations of this model can use grid resolutions coarser, 10 m and greater, than the more physics-based particle and boundary fuel models. The level set model can be used for fire spread in the surface vegetation along with the particle model for fire behavior in raised vegetation. 
\end{enumerate}
The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels and the fire spread rate is {\em predicted} by the model. The Level Set Method relies on a set of predetermined spread rates for different types of vegetation and wind speeds. The pyrolysis model, developed specifically for vegetation~\cite{Porterie:2006,Morvan:CF2004,Houssami:2016}, consists of three reactions:

@@

In FDS, the level set spread model can be implemented in the following ways:
\begin{enumerate}
\item Only the level set simulation is performed, with a constant and uniform specified wind and slope. The wind is not affected by the terrain, and there is no fire.
\item The wind field is established over the terrain, but it is ``frozen'' when the fire ignites.
\item The wind field follows the terrain, but there is no actual fire in the simulation, just front-tracking.  The level set evolves continuously in time with the flow field.
\item The wind and fire are fully-coupled and the resulting wind values are used in the head fire spread rate formula. When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
\item The wind and fire are fully-coupled in the gas-phase, but the head fire spread rate is not influenced by the wind speed. 
\end{enumerate}
When fully coupled to the CFD model, the level set function acts as an igniter as it reaches a new surface cell. The cell burns for a pre-determined amount of time. For cases where the depth of the fireline is less than the size of a grid cell, the burning time is extended while the burning rate is decreased so as to maintain conservation of fuel mass.

@@

\begin{table}[ht]
\begin{center}
\caption[Measured properties for the CSIRO Grassland Fire cases]{Measured properties for the CSIRO Grassland Fire cases~\cite{Cheney:IJWF1993}.}
\label{Properties_Grasses}
\begin{tabular}{|l|c|c|c|}
\hline
Property                        & Units        & Case C064 & Case F19    \\ \hline \hline
Wind Speed                      & m/s          & 4.6       & 4.8         \\ \hline
Ambient Temperature             & $^\circ$C    & 32.       & 34          \\ \hline
Surface Area to Volume Ratio    & m$^{-1}$     & 9770      & 12240       \\ \hline
Grass Height                    & m            & 0.21      & 0.51        \\ \hline
Bulk Mass per Unit Area         & kg/m$^2$     & 0.283     & 0.313       \\ \hline
Moisture Fraction               & \%           & 6.3       & 5.8         \\ \hline
Measured RoS                    & m/s          & 1.2       & 1.5         \\ \hline
Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) 
                                & m/s          & 1.1, 1.2, 1.2   & 1.4, 1.3, 1.4          \\ \hline
Calc'd RoS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) 
                                & m/s          & 1.3, 1.3, 1.3   & 1.5, 1.4, 1.7          \\ \hline
Calc'd RoS, Level Set Method (5 m, 10 m, 20 m resolution) 
                                & m/s          & 0.5, 0.6, 0.6   & 1.0, 1.1, 1.2          \\ \hline
\end{tabular}
\end{center}
\end{table}
